---
title: "Clustering Cycling performance data"
author: "AlexandeR Noll"
output:
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}
library(tidyverse)
library(cycleRtools)

knitr::opts_chunk$set(
  warnings = FALSE,
  message =FALSE
)
```

# Goals

The goal of this mini-project is to cluster some cycling rides: the input data is a time series that contains my wattage at every point in time.

# Data

The data was recorded using the Wahoo Kickr Snap device attached to a road bike. The `.fit` files are available on [GitHub](https://github.com/NOllAl/ClusterCycling). This file format can be read with the [`cycleRtools`]()https://cran.r-project.org/web/packages/cycleRtools/vignettes/cycling_data_analysis.html package. Let us look at one example ride:

```{r load-data}
last_ride <- read_ride("../data/2017-09-24-15-11-48.fit")

last_ride %>% 
  ggplot(aes(timer.s, power.smooth.W)) +
  geom_line()
```

The plot above has `r nrow(last_ride)` observations and corresponds to a race which lasted about 1.5 hrs. Next, we need to parse all 130 files. 

```{r pasrse-data}
parse_ride <- function(file) {
  file <- stringr::str_c("../data/", file)
  datetime <- stringr::str_sub(file, 9, 27)
  
  try({
    ride <- read_ride(file) %>% 
      mutate(datetime = datetime) %>% 
      select(datetime, timer.s, power.smooth.W)
    
    return(ride)
  })
  return(NULL)
}

load_rides <- function() {
  all_files <- list.files("../data")
  
  map_df(all_files, parse_ride)
}

all_rides <- load_rides()
```

We have `r nrow(distinct(all_rides, datetime))` rides an totally `r nrow(all_rides)` observations.

The data are downsampled to minutewise to make the clustering faster.

```{r nested-ts}
all_rides_nested <- all_rides %>% 
  group_by(datetime, minute = floor(timer.s / 60)) %>% 
  summarise(power.smooth.W = mean(power.smooth.W)) %>% 
  group_by(datetime) %>% 
  nest()
```

Now, we use the `dtwclust` package to combine DTW with a clustering technique. This combination uses DTW to compute pairwise distances and then applies a clustering algorithm to the resulting distance matrix (obiously, the clustering algorithm needs to be able to work with a distance matrix only).

```{r dtw-clust}
library(dtwclust)
set.seed(1)
clustering <- all_rides_nested$data %>% map(pull, power.smooth.W) %>% 
  tsclust(k = 3, type = "partitional")
```

Let us make a plot:

```{r plot-clustering}
all_rides_nested %>% 
  mutate(cluster = factor(clustering@cluster)) %>% 
  unnest() %>% 
  ggplot(aes(minute, power.smooth.W, color = cluster)) +
  geom_line(aes(group = datetime), alpha = 0.10) +
  stat_summary(fun.y = "mean", aes(group = cluster, color = cluster), geom = "line") +
  coord_cartesian(xlim = c(0, 90))
```

I tried a couple of cluster numbers and partition types (no optimization). The results above look quite interesting:

1. Cluster 1 (red) contains interval trainings, i.e. high efforts followed by low efforts.
2. Cluster 3 (blue) correspoinds to *races*. This starts out with relatively high efforts and they stay high.
3. Cluster 2 (green) are recovery rides.

There is much more that could be done (optimization, e.g.).

